Constraining gravitational wave amplitude birefringence with GWTC-3

### 2022-11-25 src/tex/ms.tex
update

\section{Method}
\label{sec:Method}
% Waveform modification
GW consist of two linear polarizations (i.e. $+$ and $\times$) similar to
electromagnetic waves, which could be transformed into two circular
polarizations (i.e. left-handed and right-handed) by
\begin{equation}
    h_{\mathrm{L/R}} = \frac{h_+ \pm i h_\times}{\sqrt{2}}\,,
\end{equation}
where $h_{\mathrm{L/R}}$ are the Fourier amplitude of the left-handed and
right-handed polarization of the waveform, $h_+$ and $h_\times$ are the Fourier
amplitude of the plus and cross polarization of the waveform. According to
\textbf{[modification reference]}, we modified the waveform model by
\begin{equation}
    h_\mathrm{L/R}^{\mathrm{br}}=
    h_\mathrm{L/R}^{\mathrm{GR}}\times
    \exp\left(\pm\kappa\frac{d_C}{1\mathrm{Gpc}}\frac{f}{100\mathrm{Hz}}\right)\,,
\end{equation}
where $h_\mathrm{L/R}^{\mathrm{br}}$ is the Fourier amplitude of the modified
waveform with amplitude birefringence, $h_\mathrm{L/R}^{\mathrm{GR}}$ is the
Fourier amplitude of the waveform at the source, $\kappa$ is the dimensionless
opacity parameter that represent the strength of the birefringence, $d_C$ is the
comoving distance to the source, and $f$ is the Fourier frequency of the
waveform. We assume GWs are generated by the binary as GR predicts, as the
modification at the source is small before the effect accumulate during the
propagation. As the GWs propagate, the effect of birefringence will be built up
with the number of cycles which depends on the distance traveled and the
frequency of the GWs.

% Effect on inclination PE
This modification changes the ratio of the left-handed and right-handed polarization of the waveform, which affects the PE on the inclination as well.
In GR, the amplitude ratio of the left-handed and right-handed polarizations only depends on the inclination for a nonprecessing binary.
The relationship between the amplitude ratio and the inclination is
\begin{equation}
    \frac{h_\mathrm{L}}{h_\mathrm{R}}=\left(\frac{1-\cos\iota}{1+\cos\iota}\right)^2\,,
\end{equation}
where $h_\mathrm{L}$ and $h_\mathrm{R}$ are the Fourier amplitude of left-handed and right-handed polarizations of the GWs and $\iota$ is the inclination, the angle between our line of sight and the orbital angular momentum of the binary.
PE on $\iota$ takes the amplitude ratio into account. Therefore, the PE on $\iota$ depends on the amplitude ratio.

@@

For some previous studies on the birefringence property such as \citet{Okounkova_2022}, they assume the effect of birefringence is independent of the frequency, which is a zeroth-order approximation of the birefringence model in Chern-Simons gravity.
This would create a degeneracy between $\kappa$ and $\iota$, as they can affect the amplitude ratio in the same way.
To reconstruct the amplitude ratio from the interferometer data, an $\iota$ representing a more face-off inspiral can pair with a positive $\kappa$, or an $\iota$ representing a more face-on inspiral with a negative $\kappa$.
As a result, many pairs of $\iota$ and $\kappa$ can be plausible with the frequency independence.

>>>>

\section{Method}
\label{sec:Method}
% Waveform modification
GW consist of two linear polarizations (i.e. $+$ and $\times$) similar to
electromagnetic waves, which could be transformed into two circular
polarizations (i.e. left-handed and right-handed) \kw{GW consist of two
independent polarization modes, which we can choose to represent them in either
the linear or circular basis. Better rewrite this sentence.}by
\begin{equation}
    h_{\mathrm{L/R}} = \frac{h_+ \pm i h_\times}{\sqrt{2}}\,,
\end{equation}
where $h_{\mathrm{L/R}}$ are the Fourier amplitude \kw{The equation above is not
specifically in the Fourier space, right?} of the left-handed and right-handed
polarization of the waveform, $h_+$ and $h_\times$ are the Fourier amplitude of
the plus and cross polarization of the waveform. According to
\textbf{[modification reference]}, we modified the waveform model by
\begin{equation}
    h_\mathrm{L/R}^{\mathrm{br}}=
    h_\mathrm{L/R}^{\mathrm{GR}}\times
    \exp\left(\pm\kappa\frac{d_C}{1\mathrm{Gpc}}\frac{f}{100\mathrm{Hz}}\right)\,,
\end{equation}
where $h_\mathrm{L/R}^{\mathrm{br}}$ is the Fourier amplitude of the modified
waveform with amplitude birefringence, $h_\mathrm{L/R}^{\mathrm{GR}}$ is the
Fourier amplitude of the waveform at the source \kw{Just say this is the
amplitude in the GR case.}, $\kappa$ is the dimensionless opacity parameter that
represent the strength of the birefringence, $d_C$ is the comoving distance to
the source, and $f$ is the Fourier frequency of the waveform. We assume GWs are
generated by the binary as GR predicts, as the modification at the source is
small before the effect accumulate during the propagation. As the GWs propagate,
the effect of birefringence will be built up with the number of cycles which
depends on the distance traveled and the frequency of the GWs.

% Effect on inclination PE
This modification changes the ratio of the left-handed and right-handed polarization of the waveform, which affects the PE on the inclination as well.
In GR, the amplitude ratio of the left-handed and right-handed polarizations only depends on the inclination for a nonprecessing binary. \kw{Swap the order of the two sentences.}
The relationship between the amplitude ratio and the inclination is
\begin{equation}
    \frac{h_\mathrm{L}}{h_\mathrm{R}}=\left(\frac{1-\cos\iota}{1+\cos\iota}\right)^2\,,
\end{equation}
where $h_\mathrm{L}$ and $h_\mathrm{R}$ are the Fourier amplitude of left-handed
and right-handed polarizations of the GWs \kw{You don't have to redefine the
symbols everytime. Just do it once when you introduce the symbol.} and $\iota$
is the inclination, the angle between our line of sight and the orbital angular
momentum of the binary. PE on $\iota$ takes the amplitude ratio into account. \kw{You don't need the sentence in front of this comment.}
Therefore, the PE on $\iota$ depends on the amplitude ratio.

@@

For some previous studies on the birefringence property such as \citet{Okounkova_2022} \kw{In \citet{Okounkova_2022},}, they assume the effect of birefringence is independent of the frequency, which is a zeroth-order approximation of the birefringence model in Chern-Simons gravity.
This would create a degeneracy between $\kappa$ and $\iota$, as they can affect the amplitude ratio in the same way.
To reconstruct the amplitude ratio from the interferometer data, an $\iota$ representing a more face-off inspiral can pair with a positive $\kappa$, or an $\iota$ representing a more face-on inspiral with a negative $\kappa$.
As a result, many pairs of $\iota$ and $\kappa$ can be plausible with the frequency independence.
